Enhancing Adversarial Robustness for Deep Metric Learning

### 2021-10-24 robdml.tex
tex

\textbf{Adversarial Attack.}
%
Szegedy \etal~\cite{l-bfgs} find misclassification of DNN can be triggered by
an imperceptible adversarial perturbation to the input image.
%
Ian \etal~\cite{fgsm} speculate the reason is that DNN being locally linear
with respect to the adversarial perturbation.
%
A series of following works attempt to more efficiently compromise the DNNs,
such as BIM~\cite{i-fgsm}, C\&W~\cite{cw}, PGD~\cite{madry}, and
APGD~\cite{apgd}.
%
These attacks are based on the first-order gradient of the classification loss
with respect to the input image, and hence rely on the white-box assumption
that the model architecture and parameters are accessible to the attacker,
which is impractical for real-world attacks.
%
Subsequently, black-box attack methods are developed into two types:
transferrability-based attacks and query-based attacks.
%
Transferrability-based attacks~\cite{di-fgsm,universal} relies on the
observation that image-agnostic (will take effect when applied to any image)
and model-agnostic (will take effect when applied to any model) adversarial
perturbation are possible.
%
Query-based attacks~\cite{nes-atk,spsa-atk} only need the logit score output or
the label output from a classifier, based on which the gradient could be
estimated from the response to repetitive queries in order to figure out the
adversarial perturbation.

@@

\textbf{Deep Metric Learning.}
%
A wide range of application problem such as image retrieval~\cite{imagesim2},
cross-modal retrieval~\cite{ladderloss}, and face recognition~\cite{facenet}
can be formularized as a deep metric learning problem.
%
Deep metric learning has been found vulnerable to adversarial attacks as
well~\cite{advrank,advorder}, which will result in undesired implications on
safety, security, or fairness of a deep metric learning application.
%
In contrast, the defense methods for enhancing the adversarial robustness of
deep metric learning are less explored.
%
Zhou \etal~\cite{advrank} present an adversarial training method with
adversarial examples maximizing the embedding move distance off its original
location.
%
Anti-Collapse Triplet~\cite{robrank} forces the model to separate collapsed
positive and negative samples in order to learn robust features.
%
However, the existing defense methods refrain from adopting Madry's min-max
adversarial training paradigm due to the model being too easy to collapse with
excessively hard adversarial samples.

@@

Similar to what have been found for the classifiers, recent
works~\cite{robrank,advrank,advorder} also suggest that the embedding function
$\phi(\cdot)$ in DML is vulnerable to adversarial attacks.
%
Specifically, an imperceptible adversarial perturbation $r$ is added to the
input image $X$ ($\|r\|_p \leq \varepsilon$, $X+r\in \mathcal{X}$), so that its
embedding vector $\tilde{\vx}=\phi(X+r)$ will be moved off its original
location towards other positions to fulfill the attacker's goal.

Although attacks against DML has been widely studied~\cite{advrank,advorder},
the defense methods for improving adversarial robustness is much less explored.
%
As adversarial training~\cite{madry} remains to be one of the most effective
defense for classification, existing defense methods for DML are also
adversarial training-based.
%
Embedding-Shifted Triplet (EST)~\cite{advrank} adopts adversarial counterpars
of $\va,\vp,\vn$ with maximum embedding move distance off their original
locations, \ie,
$L_\text{EST}=L_\text{T}(\tilde{\va},\tilde{\vp},\tilde{\vn};\gamma)$ where
$\tilde{\va}=\phi(X+r^*)$, $r^*=\arg\max_{r}d_\phi(X+r, X)$.
%
Anti-Collapse Triplet (ACT)~\cite{robrank} collapses the embedding vectors of
positive and negative sample, and enforces the model to separate them apart,
\ie, $L_\text{ACT}=L_\text{T}(\va, \overrightarrow{\vp},
\overleftarrow{\vn};\gamma)$, where $[\overrightarrow{\vp},\overleftarrow{\vn}]
=[\phi(X_p+r_p^*), \phi(X_n+r_n^*)]$, and $[r_p^*,r_n^*]=\arg\min_{r_p,r_n}
d_\phi(X_p+r_p, X_n+r_n)$.
%
However, compared to the standard min-max adversarial training
paradigm~\cite{madry}, these methods merely indirectly increase the loss value,
and thus, suffer from inefficient learning because the adversary is not strong
enough.
%
\oo{Furthermore, min-max form -- free adversarial training}.

@@

>>>>

\textbf{Adversarial Attack.}
%
Szegedy \etal~\cite{l-bfgs} find misclassification of DNN can be triggered by
an imperceptible adversarial perturbation to the input image.
%
Ian \etal~\cite{fgsm} speculate the reason is that DNN being locally linear
with respect to the adversarial perturbation.
%
Subsequent studies can compromise the DNNs more effectively, such as
BIM~\cite{i-fgsm}, PGD~\cite{madry}, and APGD~\cite{apgd}, which are
unanimously based on the first-order gradient of the loss, and rely on the
white-box assumption of fully accessible model details.
%
In contrast, query-based black-box attacks~\cite{nes-atk,spsa-atk} are more
practical for real-world scenarios, as they requires the least amount of
information from the model for estimating the gradients.

@@

\textbf{Deep Metric Learning DML.}
%
A wide range of application problem such as image retrieval~\cite{imagesim2},
cross-modal retrieval~\cite{ladderloss}, and face recognition~\cite{facenet}
can be formularized as a deep metric learning problem.
%
Deep metric learning has been found vulnerable to adversarial attacks as
well~\cite{advrank,advorder}, which will result in undesired implications on
safety, security, or fairness of a deep metric learning application.
%
In contrast, the defense methods for enhancing the adversarial robustness of
deep metric learning are less explored.
%
Zhou \etal~\cite{advrank} present an adversarial training method with
adversarial examples maximizing the embedding move distance off its original
location.
%
Anti-Collapse Triplet~\cite{robrank} forces the model to separate collapsed
positive and negative samples in order to learn robust features.
%
However, the existing defense methods refrain from adopting Madry's min-max
adversarial training paradigm due to the model being too easy to collapse with
excessively hard adversarial samples.

\textbf{Attacks in DML.}
%
Similar to what have been found for the classifiers, recent
works~\cite{robrank,advrank,advorder} also suggest that the embedding function
$\phi(\cdot)$ in DML is vulnerable to adversarial attacks.
%
Specifically, an imperceptible adversarial perturbation $r$ is added to the
input image $X$ ($\|r\|_p \leq \varepsilon$, $X+r\in \mathcal{X}$), so that its
embedding vector $\tilde{\vx}=\phi(X+r)$ will be moved off its original
location towards other positions to fulfill the attacker's goal.

\textbf{Defenses in DML.}
%
Although attacks against DML has been widely studied~\cite{advrank,advorder},
the defense methods for improving adversarial robustness is much less explored.
%
As adversarial training~\cite{madry} remains to be one of the most effective
defense for classification, existing defense methods for DML are also
adversarial training-based.
%
Embedding-Shifted Triplet (EST)~\cite{advrank} adopts adversarial counterpars
of $\va,\vp,\vn$ with maximum embedding move distance off their original
locations, \ie,
$L_\text{EST}=L_\text{T}(\tilde{\va},\tilde{\vp},\tilde{\vn};\gamma)$ where
$\tilde{\va}=\phi(X+r^*)$, $r^*=\arg\max_{r}d_\phi(X+r, X)$.
%
Anti-Collapse Triplet (ACT)~\cite{robrank} collapses the embedding vectors of
positive and negative sample, and enforces the model to separate them apart,
\ie, $L_\text{ACT}=L_\text{T}(\va, \overrightarrow{\vp},
\overleftarrow{\vn};\gamma)$, where $[\overrightarrow{\vp},\overleftarrow{\vn}]
=[\phi(X_p+r_p^*), \phi(X_n+r_n^*)]$, and $[r_p^*,r_n^*]=\arg\min_{r_p,r_n}
d_\phi(X_p+r_p, X_n+r_n)$.
%
However, compared to the standard min-max adversarial training
paradigm~\cite{madry}, these methods merely indirectly increase the loss value,
and thus, suffer from inefficient learning because the adversary is not strong
enough.
%
\oo{Furthermore, min-max form -- free adversarial training}.

@@

@@

\begin{equation}
	L_\text{ICS} = \lambda \cdot L_\text{T}(
	\va, \tilde{\va}, \vp; 0)
\end{equation}
